Windowed Developer Topic Analysis

### 2009-04-08 lda-paper.tex
More lda paper changes from ric's suggestions

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is where concepts are automatically extracted from source code or
documents related to the source code.  Concept location is how high
level concepts relate to low level entities like source code, for
instance when fixing a bug how do the concepts in the bug report map
back to the source code?  Semantic Clustering has also been used for
similar reasons~\cite{1698774,1566153} as it is similar to LSI.

Grant et al.~\cite{scottcordy} have used an alternative technique to
LSI and LDA, they used Independent Component Analysis to separate
topic signals from software code.

Our technique differs in the sense we apply LDA temporally and we
apply to change comments.

@@

In our first exploratory pass we wanted to see

Looking at 30 day non-overlapping windows of revisions for MySQL 3.23
(20 topics, top 10 words per topic) we found there were common words
across topic clusters such as diffs, annotate and history. There were
notable transitional topics such as in the first window the word
``bitkeeper'' appears (probably when they adopted Bitkeeper for their
source control system) yet in the following windows there were no
mentions of Bitkeeper. ``RENAME'' also only appeared once in the first
window and never again.

A subsampling of the notable topic words is displayed in Table
\ref{tab:portability}. We tracked one topic cluster which seemed
related to portability. We picked out terms which seem to summarize
those topics.

%Often words are shared across topics, see table \ref{tab:portability}
%for tracking topics about portability.

After viewing these topic is was obvious that we would have to combine
try to track the evolution of topics over the windows or rank topics
by similarity.

@@

\begin{table*}
\centering
\begin{tabular}{|ccc|l|}
\hline
2000 &  Jul &  31 &    chmod \\
2000 &  Sep &  29 &    fixes benchmark logging windows \\
2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
\hline
2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\

2001 &  Aug &  25 &    tables row version [a] \\
\hline
2001 &  Sep &  24 &    update checksum merge \\
2001 &  Oct &  24 &    fixed fix \\
2001 &  Dec &  23 &    HPUX SCO fix \\
\hline
2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
2002 &  Mar &  23 &    small buf fix [b]  \\
\hline
2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &  18 &    HPUX11 compiler HP \\
\hline
2003 &  Feb &  16 &    Linux errno  [c] \\
2003 &  Mar &  18 &    alarm bookmark bug [c] \\
\hline
2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Tracking topics associated with the word portability, note some continuous blocks}
\label{tab:portability}
\end{table*}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends. The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time. Longer topics are topics which reoccur in adjacent window.}
  \label{fig:lda}
\end{figure*}

>>>>

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
Concept analysis aims to automatically extract concepts from source
code or documents related to the source code.  Concept location is how
high level concepts relate to low level entities like source code, for
instance when fixing a bug how and where do the concepts in the bug
report map back to the source code?  Semantic Clustering has also been
used for similar reasons~\cite{1698774,1566153} as it is similar to
LSI.

Grant et al.~\cite{scottcordy} have used an alternative technique,
rather than LSI and LDA, called Independent Component Analysis to
separate topic signals from software code.

Our technique differ from the previous techniques because we apply LDA
locally to month-long windows of commit comments, where as other
approaches apply LDA once to the entire project.

@@

In our first exploratory pass we wanted to see if LDA could provide
interesting topics extracted from a real system. We took the
repository of MySQL 3.23, extracted the commits, grouped them by 30
day non-overlapping windows then applied LDA to each of these
windows. We asked LDA to make 20 topics per window and we would pay
attention to the top 10 most frequent words in that topic.  We found
there were common words across topic clusters, such as diffs, annotate
and history, these words probably should be viewed as stop
words. There were notable transitional topics such as in the first
window the word ``bitkeeper'' appears (probably MySQL they adopted
Bitkeeper for their source control system) yet in the following
windows there were no mentions of Bitkeeper. ``RENAME'' also only
appeared once in the first window and never again. Per each topic we
tried to name the purpose of the topic, we found that with little
familarity with the code base that naming the topic was not difficult.

A subsampling of the notable topic words is displayed in Table
\ref{tab:portability}.  To name the topics we picked out terms which
seem to summarize those topics.  After viewing these topic is was
obvious that we would have to combine try to track the evolution of
topics over the windows or rank topics by similarity, thus if they
were similar enough we joined them into a trend. Figure \ref{fig:lda}
displays a sample plot of the extracted topics in table
\ref{tab:portability}.

@@

% \begin{table*}
% \centering
% \begin{tabular}{|ccc|l|}
% \hline
% 2000 &  Jul &  31 &    chmod \\
% 2000 &  Sep &  29 &    fixes benchmark logging windows \\
% 2000 &  Nov &  28 &    typo fix insert\_multi\_value \\
% 2001 &  Jan &  27 &    fixes Innobase Cleanups auto-union \\
% 2001 &  Mar &  28 &    2 topics bugfix, logging , TEMPORARY,  \\
% \hline
% 2001 &  Jul &  26 &    update Allow TABLES LOCK [a] \\

% 2001 &  Aug &  25 &    tables row version [a] \\
% \hline
% 2001 &  Sep &  24 &    update checksum merge \\
% 2001 &  Oct &  24 &    fixed fix \\
% 2001 &  Dec &  23 &    HPUX SCO fix \\
% \hline
% 2002 &  Feb &  21 &    net buffer length  max\_allowed\_packet [b] \\
% 2002 &  Mar &  23 &    small buf fix [b]  \\
% \hline
% 2002 &  May &  22 &    [popular] fix SCO OSF1 table\_name \\
% 2002 &  Nov &  18 &    HPUX11 compiler HP \\
% \hline
% 2003 &  Feb &  16 &    Linux errno  [c] \\
% 2003 &  Mar &  18 &    alarm bookmark bug [c] \\
% \hline
% 2003 &  Sep &  14 &    Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
% \hline
% \end{tabular}
% \caption{Tracking topics associated with the word portability, note some continuous blocks}
% \label{tab:portability}
% \end{table*}

\begin{table*}
\centering
\begin{tabular}{|cc|l|}
\hline
2000 &  Jul &      chmod \\
2000 &  Sep &      fixes benchmark logging Win32 \\
2000 &  Nov &      fixes insert\_multi\_value \\
2001 &  Jan &      fixes Innobase Cleanups auto-union \\
2001 &  Mar &      bugfix logging  TEMPORARY  \\
\hline         
2001 &  Jul &      TABLES update Allow LOCK \\

2001 &  Aug &      TABLES row version \\
\hline         
2001 &  Sep &      update checksum merge \\
2001 &  Oct &      fixed fix \\
2001 &  Dec &      HPUX SCO fix \\
\hline         
2002 &  Feb &      net buffer length  max\_allowed\_packet \\
2002 &  Mar &      small buf fix  \\
\hline         
2002 &  May &      [popular] fix SCO OSF1 table\_name \\
2002 &  Nov &      HPUX11 compiler HP \\
\hline         
2003 &  Feb &      Linux errno   \\
2003 &  Mar &      alarm bookmark bug \\
\hline         
2003 &  Sep &      Auto logging merge windows distribution fix 64-bit 4.0 Cleanup \\
\hline
\end{tabular}
\caption{Sub-sampled topics from MySQL 3.23, some with continuous topics, these tokens were pulled from LDA topic. Each token is a summary of one LDA generated topic from MySQL 3.23 commit logs.}
\label{tab:portability}
\end{table*}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the kind of plot we eventually want to produce: named topics and topic trends. The horizontal axis is time in months and the vertical axis used to stack topics that occur at the same time. Longer topics are topics which reoccur in adjacent window. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}
